Robust Bayesian Analysis of Causal Inference Problems

### 2022-10-31 JASA-template.tex
bayes

This allows us to define intermediate latent variables 
\begin{equation}
	T_i^* = \gamma_0 + X_i\gamma +u_i
\end{equation}
where, $u_i\sim\mathcal{N}(0,1)$. Therefore, $T_i=1$ if $T_i^*>0$ and
$T_i=0$ if $T_i^*\le0$.

Now, following the approach of \citet{koch2018}, we define an adjusted
output vector $W\coloneqq(Y, T^*)^T$ and corresponding $2n\times(2p+1)$ dimensional design matrix
\begin{align}
    Z &=
    \begin{bmatrix}
     X_O & 0 \\
     0 & X
    \end{bmatrix},
\end{align}
where, $X_O = [T, X]$. Then, assuming Gaussian error term, we have the following likelihood function
\begin{align}
W\mid Z, \alpha, \beta, \gamma, \sigma^2 \sim\normal\left(Z\nu, \sigma^2\mathbf{I}_{2n}\right),
\end{align}
where $\nu = (\alpha, \beta, \gamma)^T$.

\section{Bayesian Hierarchical model}\label{sec:bayes}
\subsection{Hierarchical model}

\begin{figure}
	\centering
	\begin{tikzpicture}[params/.style={circle, draw=black!60, very thick, minimum size=7mm},
	hyper/.style={circle, draw=black!60, fill=black!20, thick, minimum size=7mm},
	latent/.style={rectangle, draw=black!60, fill=black!10, dashed, minimum size=7mm},
	data/.style={rectangle, draw=black!60, thick, minimum size=8mm}]
	\node[params] (1) at (0,0) {$\pi$};
	\node[data] (2) at (3,0) {$X$};
	\node[data] (3) at (6,0) {$T$};
	\node[params] (4) at (1.5,1.5) {$\gamma$};
	\node[latent] (5) at (4.5,1.5) {$T^*$};
	\node[params] (6) at (1.5,-1.5) {$\beta$};
	\node[data] (7) at (3.5,-1.5) {$Y$};
	\node[hyper] (8) at (-1.5,-3) {$\sigma^2$};
	\node[params] (9) at (3,-3) {$\beta_0$};
	\node[params] (10) at (5.5,-1.5) {$\beta_T$};
	\node[hyper] (11) at (-1.5,-1) {$s$};
	\node[hyper] (12) at (-1.5,1) {$\alpha$};
	\draw[black, dashed] (-0.75,-3.75) rectangle (7,2.5);

	\path (1) edge[->]  (6);
	\path (1) edge[->]  (4);
	\path (8) edge[->]  (6);
	\path (6) edge[->]  (7);
	\path (2) edge[->]  (7);
	\path (2) edge[->]  (5);
	\path (5) edge[<-] (4);
	\path (5) edge[->] (3);
	\path (3) edge[->]  (7);
	\path (9) edge[->]  (7);
	\path (10) edge[->]  (7);
	\path (2) edge[dashed][->] (3);
	\path (8) edge[bend right = 60][->]  (10);
	\path (8) edge[->]  (9);
	\path (11) edge[->]  (1);
	\path (12) edge[->]  (1);

@@

We perform a robust Bayesian analysis to obtain robust estimates. To do so, we propose the
following spike and slab priors \cite{ishwaran2005} to specify $\beta$ and $\gamma$.
For $1\le j\le p$,
\begin{align}
    \beta_j \mid \pi_{\beta}, \sigma^2 &\sim 
    \pi_{\beta}\normal\left(0, \tau_1^2\sigma^2\right) 
    + (1-\pi_{\beta}) \normal\left(0, \tau_0^2\sigma^2\right)\\
    \gamma_j \mid \pi_{\gamma} &\sim 
    \pi_{\gamma}\normal\left(0, \tau_1^2\right) 
    + (1-\pi_{\gamma}) \normal\left(0, \tau_0^2\right)\\
    \alpha\mid \sigma^2 &\sim \normal(0, \sigma^2)\\
    \sigma^2&\sim \text{InvGamma}(a, b)\\
    \pi_{\beta}, \pi_{\gamma}&\sim\text{Beta}\left(sq, s(1-q)\right).
\end{align}
We fix sufficiently small $\tau^2_0$
$(1\gg\tau_0^2>0)$ so that  $(\beta_j, \gamma_j) = (0,0)$ has its probability mass 
concentrated around zero. Therefore, this represents the spike component of our prior specification. 
To construct the slab component, we consider $\tau_1^2$ to be large so that 
$\tau_1^2\gg\tau_0^2$. This allows the prior for $(\beta_j, \gamma_j)\not=(0,0)$ to be flat.

We use a set beta priors to specify the selection probability
$\pi_j$ of the $j$-th group where  $q_j$ represents our prior expectation of the 
selection probability ($\pi_j$) and `$s$' represents a concentration parameter. We perform our
robust Bayesian analysis on $q\coloneqq(q_1$, \dots, $q_p)\in\mathcal{P}$, where
\begin{equation}
    \mathcal{P} \coloneqq \mathcal{P}_1\times\cdots\times\mathcal{P}_p\subseteq \left(0, 1\right)^{p}.
\end{equation}

@@

>>>>

This allows us to define intermediate latent variables as suggested
by \citet{albert93}
\begin{equation}
	T_i^* = \gamma_0 + X_i\gamma +u_i
\end{equation}
where, $u_i\sim\mathcal{N}(0,1)$. Therefore, $T_i=1$ if $T_i^*>0$ and
$T_i=0$ if $T_i^*\le0$.

Now, following the approach of \citet{koch2018}, we define an adjusted
output vector $W\coloneqq(Y, T^*)^T$ and corresponding $2n\times(2p+3)$ dimensional design matrix
\begin{align}
    Z &=
    \begin{bmatrix}
     X_O & 0 \\
     0 & X_T
    \end{bmatrix},
\end{align}
where, $X_O = [T, 1_n, X]$ and $X_T = [1_n, X]$. Then, assuming a
Gaussian error term, we have the following likelihood distribution
\begin{align}
W\mid Z, \alpha, \beta, \gamma, \sigma^2 \sim\normal\left(Z\nu, \Sigma\right)\label{eq:like:group},
\end{align}
where $\nu = (\beta_T, \beta_0, \beta, \gamma_0, \gamma)^T$ and
\begin{align}
\Sigma &=
\begin{bmatrix}
\sigma^2{I}_n & 0 \\
0 & {I}_n
\end{bmatrix}.
\end{align}

\section{Bayesian Causal Estimation}\label{sec:bayes}

The likelihood formation given by \cref{eq:like:group} gives us
a foundation for Bayesian group LASSO 
\cite{xu2015} type model and look into the posterior selection
probability associated with the $j$-th predictor. There are several
ways to construct spike and slab priors which achieve 
variable selection. In our case, we consider a continuous type
\cite{ishwaran2005} prior for faster posterior
computation.

\subsection{Hierarchical model}

Let, $\pi_j$ denote the prior probability that the $j$-th
predictor is not associated with either the outcome or the 
treatment. That is, 
\begin{equation}
	\pi_j = P\left((\beta_j,\gamma_j)=(0,0)\right).
\end{equation}
Then we can define a spike and slab group LASSO so that:
for $1\le j\le p$,
\begin{align}
(\beta_j,\gamma_j)^T \mid \pi_{j}, \sigma^2 &\sim 
(1-\pi_{j})\normal\left( (0,0)^T, 
\tau_1^2\begin{bmatrix}
\sigma^2 & 0 \\
0 & 1
\end{bmatrix}\right)
+ \pi_{j} \normal\left(0, 
\tau_0^2\begin{bmatrix}
\sigma^2 & 0 \\
0 & 1
\end{bmatrix}\right)\\
%\beta_T, \beta_0\mid \sigma^2 &\sim \normal(0, \sigma^2)\\
%\gamma_0 &\sim \normal(0,1)\\
\sigma^2&\sim \text{InvGamma}(a, b)\\
\pi_{j} &\sim\text{Beta}\left(sq_j, s(1-q_j)\right).
\end{align}

We fix sufficiently small $\tau^2_0$
$(1\gg\tau_0^2>0)$ so that  $(\beta_j, \gamma_j) = (0,0)$ has its probability mass 
concentrated around zero. Therefore, this represents the spike component of our prior specification. 
To construct the slab component, we consider $\tau_1^2$ to be large so that 
$\tau_1^2\ge 1$. This allows the prior for $(\beta_j, \gamma_j)\not=(0,0)$ to be flat.

As mentioned earlier, $\pi_j$ is used as the rejection probability
of the $j$-th predictor and we use a set beta priors to specify 
these rejection probabilities where  $q_j$ represents our prior expectation of the rejection probability ($\pi_j$) and `$s$' acts as 
a concentration parameter.
For the intercept terms of the outcome model and the causal effect, 
we consider a sufficiently flat prior so that 
$\beta_0. \beta_T\sim \normal(0,\sigma^2)$. Similarly, for the
intercept term in the treatment model, we consider 
$\gamma_0\sim \normal(0,1)$.

\begin{figure}
	\centering
	\begin{tikzpicture}[params/.style={circle, draw=black!60, very thick, minimum size=7mm},
	hyper/.style={circle, draw=black!60, fill=black!20, thick, minimum size=7mm},
	post/.style={circle, draw=black!60, fill=green!20, thick, minimum size=7mm},
	latent/.style={rectangle, draw=black!60, fill=black!10, dashed, minimum size=7mm},
	data/.style={rectangle, draw=black!60, thick, minimum size=8mm}]
	\node[params] (1) at (0,0) {$\pi$};
	\node[data] (2) at (3,0) {$X$};
	\node[data] (3) at (6,0) {$T$};
	\node[params] (4) at (1.5,1.5) {$\gamma$};
	\node[latent] (5) at (4.5,1.5) {$T^*$};
	\node[params] (6) at (1.5,-1.5) {$\beta$};
	\node[data] (7) at (3.5,-1.5) {$Y$};
	\node[params] (8) at (0,-3) {$\sigma^2$};
	\node[params] (9) at (3,-3) {$\beta_0$};
	\node[params] (10) at (6,-1.5) {$\beta_T$};
	\node[hyper] (11) at (-1.5,-.8) {$s$};
	\node[hyper] (12) at (-1.5,.8) {$q$};
	\node[hyper] (13) at (-1.5,-2.2) {$a$};
	\node[hyper] (14) at (-1.5,-3.8) {$b$};
	\draw[black, dashed] (0.75,-3.7) rectangle (7,2.1);

	\path (1) edge[->]  (6);
	\path (1) edge[->]  (4);
	\path (8) edge[->]  (6);
	\path (6) edge[->]  (7);
	\path (2) edge[->]  (7);
	\path (2) edge[->]  (5);
	\path (5) edge[<-] (4);
	\path (5) edge[->] (3);
	\path (3) edge[->]  (7);
	\path (9) edge[->]  (7);
	\path (10) edge[->]  (7);
	\path (2) edge[dashed][->] (3);
	\path (8) edge[bend right = 60][->]  (10);
	\path (8) edge[->]  (9);
	\path (11) edge[->]  (1);
	\path (12) edge[->]  (1);
	\path (13) edge[->]  (8);
	\path (14) edge[->]  (8);

@@

@@

\subsection{Robust Bayesian Analysis}
We perform our
robust Bayesian analysis on $q\coloneqq(q_1$, \dots, $q_p)\in\mathcal{P}$, where
\begin{equation}
\mathcal{P} \coloneqq \mathcal{P}_1\times\cdots\times\mathcal{P}_p\subseteq \left(0, 1\right)^{p}.
\end{equation}
